From clusters to queries: exploiting uncertainty in the modularity landscape of complex networks

### 2018-04-30 article/article.tex
boilerplate

\begin{figure*}
\centering

  \begin{subfigure}[ht]{0.4\textwidth}
    \includegraphics[width=\textwidth]{images/search_space_3.png}
    \caption{Modularity optimisation space of an \textit{E. coli} metabolic network.}
    \label{fig:modular_search_space}
  \end{subfigure}
  \begin{subfigure}[ht]{0.58\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/meth_fig/fig1_desc.eps}
    \caption{Outline of the proposed approach to querying networks by using multiple, high quality representations of modular networks.}
    \label{fig:algorithm_outline}
  \end{subfigure}
 \caption{
 Modularity search space of an \textit{E. coli} metabolic network.
 Distance between partitions is calculated using variation of information \cite{meilua2003comparing} and dimensionality reduction is performed using curvilinear component analysis \cite{demartines1997curvilinear}.
  The inset (top) demonstrates the landscape of the high modularity region. 
  Figure generated with the software of Good \textit{et al.}\cite{good2010performance}.
 }
\end{figure*}

@@

\subsection{Synthetic networks}
In this section we test the method on a number benchmark networks constructed with a known, ground-truth community structure.
To evaluate how our method performs we demonstrate results against two benchmark networks the conventionally used LFR benchmark \cite{lfr} for community detection algorithms.
We test the area under the ROC curve (AUC) scores for networks synthetically generated with the LFR benchmark.
The community distribution is defined with a powerlaw coefficient of $-1.0$ and the degree distribution is defined with a powerlaw coefficient of $-2.0$.
\textbf{Non-overlapping modules.} Figure \ref{fig:auc_no_overlap} represents tests on 100 networks with different randomly selected seed nodes.
As one would expect, the prediction of the method drops off steeply where communities are less defined above a mixing coefficient of 0.6.
Using larger seed sets also improves accuracy with some prediction of true communities being possible at extremely high levels of mixing.

\begin{figure*}[ht]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_mo_overlap_auc_1000.eps}
        \caption{1000 nodes}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_mo_overlap_auc_5000.eps}
        \caption{5000 nodes}
    \end{subfigure}
    \caption{Non-overlapping community varying seed nodes on LFR networks with 1000 and 5000 nodes.
     Data points represent mean auc scores for all communities on 100 sampled networks at varying mixing coefficients.
     Error bars represent standard deviations.}
\end{figure*}

\begin{figure*}[ht]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_overlap_auc_1000.eps}
        \caption{1000 nodes}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_overlap_auc_1000.eps}
        \caption{5000 nodes}
    \end{subfigure}
    \caption{Overlapping community varying seed nodes on LFR networks with 1000 and 5000 nodes.
     Data points represent mean auc scores for all communities on 100 sampled networks at varying overlapping fraction of nodes.
     Error bars represent standard deviations.}
\end{figure*}

@@

\begin{figure*}[ht]
\includegraphics[width=\textwidth]{images/ppi_example_combined.eps}
\caption{An example of the GO term \textit{``SNAP receptor activity''} (16 nodes) on the \textit{Arabidopsis thaliana} protein-protein interaction network.
(A) Shows the full network with 3 randomly chosen query nodes from the full query set.
(B) Shows a reduced view of the network where a threshold is selected, reducing the network to 404 nodes.
Blue nodes indicate those in the randomly selected 3 query nodes, yellow the remaining nodes in the term and the shade of red relates to the strength of the relationship each node has to the query set.
(C) Shows an example ROC curve from the cross validation described in the main text. 
Based on 3 randomly selected query nodes containing this GO term.
}
\label{fig:query_example}
\end{figure*}

>>>>

\begin{figure}[ht]
    \includegraphics[width=0.45\textwidth]{images/search_space_3.png}
    \caption{Modularity search space of an \textit{E. coli} metabolic network.
 Distance between partitions is calculated using variation of information \cite{meilua2003comparing} and dimensionality reduction is performed using curvilinear component analysis \cite{demartines1997curvilinear}.
  The inset (top) demonstrates the landscape of the high modularity region. 
  Figure generated with the software of Good \textit{et al.}\cite{good2010performance}.}
    \label{fig:modular_search_space}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.45\textwidth]{images/meth_fig/fig1_desc.eps}
    \caption{Outline of the proposed approach to querying networks by using multiple, high quality representations of modular networks.}
    \label{fig:algorithm_outline}
\end{figure}

@@

\subsection{Synthetic networks}
In this section we test the method on benchmark networks constructed with a known, ground-truth community structure.
To evaluate how our method performs we use the undirected, unweighted LFR benchmark \cite{lancichinetti2008benchmark} in overlapping and non-overlapping forms.
We test the area under the ROC curve (AUC) scores for networks synthetically generated with the LFR benchmark.
In these tests, the community distribution is defined with a powerlaw coefficient of $-1.0$, the degree distribution is defined with a powerlaw coefficient of $-2.0$ and the number of nodes is stated.

\textbf{Non-overlapping modules.} Figure \ref{fig:auc_no_overlap} represents tests on 100 networks with 1000 and 5000 nodes varying the mixing coefficient (number of edges between communities).
As one would expect, the prediction of the method drops off steeply where communities are less defined above a mixing coefficient of 0.6.
Using larger seed sets also improves accuracy with some prediction of true communities being possible at extremely high levels of mixing.

\textbf{Overlapping modules.} 
In figure \ref{fig:auc_overlap} we show the results of network performance when tested against an increasing level of overlapping communities.
For these tests we fix the mixing coefficient at $0.3$.
Here, each vertex can belong to up to 4 communities.
In order to test performance we varied the fraction of nodes that are in more than one community.
Interestingly, the method still has an AUC score above 0.5 when all nodes are placed in multiple communities.
This indicates that the method is capable of uncovering latent overlapping memberships even when given a relatively small number of seed nodes.

\begin{figure*}[ht]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_mo_overlap_auc_1000.eps}
        \caption{1000 nodes}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_mo_overlap_auc_5000.eps}
        \caption{5000 nodes}
    \end{subfigure}
    \caption{Non-overlapping community varying seed nodes on LFR networks with 1000 and 5000 nodes.
     Data points represent mean auc scores for all communities on 100 sampled networks at varying mixing coefficients.
     Error bars represent standard deviations.}
     \label{fig:auc_no_overlap}
\end{figure*}

\begin{figure*}[ht]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_overlap_auc_1000.eps}
        \caption{1000 nodes}
    \end{subfigure}
    \begin{subfigure}[b]{0.48\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_overlap_auc_5000.eps}
        \caption{5000 nodes}
    \end{subfigure}
    \caption{Overlapping community varying seed nodes on LFR networks with 1000 and 5000 nodes.
     Data points represent mean auc scores for all communities on 100 sampled networks at varying overlapping fraction of nodes.
     Error bars represent standard deviations.}
     \label{fig:auc_overlap}
\end{figure*}

@@

\begin{figure*}[ht]
\includegraphics[width=\textwidth]{images/ppi_example_combined.eps}
\caption{An example of the GO term \textit{``SNAP receptor activity''} (16 nodes) on the \textit{Arabidopsis thaliana} protein-protein interaction network.
(A) Shows the full network with 3 randomly chosen query nodes from the full query set.
(B) Shows a reduced view of the network where a threshold of $\mu_i(S) > 0.8$ is selected, reducing the network to 404 nodes.
Blue nodes indicate those in the randomly selected 3 query nodes, yellow the remaining nodes in the term and the shade of red relates to the strength of the relationship each node has to the query set.
(C) Shows an example ROC curve from the cross validation described in the main text. 
Based on 3 randomly selected query nodes containing this GO term.
}
\label{fig:query_example}
\end{figure*}
